---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
```{r, message = FALSE}

libraries <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest" , "metaforest", "rstatix", "labelled","kableExtra", 'knitr', "forcats", 'ggrepel', 'writexl')
 
new.packages <- libraries[!(libraries %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(libraries, require, character.only = TRUE)
 
# _fixed dataset has removed any misalignment of subjects across sites
path = file.path( "Master_BD_080720_withPredictions_withNewHajek_fixed.sav")
df = read_sav(path)

df <- df %>% 
  mutate_if(is.character, list(~na_if(.,""))) %>%
  filter(!SubjID %in%  c('con_11',  'pat_13')) %>%
    arrange(Site) 
siteLoc <- read_excel("DescriptiveData_080720_2.xlsx", sheet=3)

# create function to replace spaces with underscores
replaceSpace <- function(dataframe) {
   gsub(x = names(dataframe), pattern = " ", replacement = "_") 
} 

names(siteLoc) <- replaceSpace(siteLoc)
names(df) <- replaceSpace(df) 

# merge the main dataset with the Site Location
df <- merge(df, siteLoc, by=c("PIName","Site"  )) 
df <- df %>% 
  filter(Age <= 75  & Age  >=18) 
BD_only <- df %>% filter(Dx==1) 
 
df %>% group_by(Dx) %>% summarize(mean = mean(BrainPAD), sd = sd(BrainPAD)) %>% round(2)
 
```


### Plot BrainPAD Lithium, Antiepileptics, SGA vs. not on each of those 
```{r}
meanBrainPAD_Li <- BD_only %>% 
  filter(!is.na(Li)) %>%
  group_by(Li) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2), count =   n())    %>%
  `colnames<-`(c("moodStabilizer", "brainPAD_mean", "count"))  %>%
  mutate(moodStabilizer =
                  case_when(
                    moodStabilizer == 0   ~ 'No Lithium',
                    moodStabilizer == 1  ~ 'On Lithium', 
                  ))

meanBrainPAD_AntiE <- BD_only %>% 
  filter( !is.na(AntiEpileptic)) %>%
  group_by(AntiEpileptic) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2),
                        count = n())    %>%
  `colnames<-`(c("moodStabilizer", "brainPAD_mean", "count"))  %>%  mutate(moodStabilizer =
                  case_when(
                    moodStabilizer == 0   ~ 'No AntiEpileptics',
                    moodStabilizer == 1  ~ 'On AntiEpileptics', 
                  )) 
 
meanBrainPAD_SGA <- BD_only %>% 
  filter( !is.na(Gen2AntiPsych)) %>%
  group_by(Gen2AntiPsych) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2),
                     count = n())    %>%
  `colnames<-`(c("moodStabilizer", "brainPAD_mean", "count"))%>%
  mutate(moodStabilizer =
                  case_when(
                    moodStabilizer == 0   ~ 'No SGA',
                    moodStabilizer == 1  ~ 'On SGA', 
                  )) 
brainPADs <-  rbind(meanBrainPAD_Li, meanBrainPAD_AntiE, meanBrainPAD_SGA)

melt_bd <- unlabelled(BD_only) %>% dplyr::select(BrainPAD, Li, AntiEpileptic, Gen2AntiPsych) %>%
  filter(!is.na(Li), !is.na(AntiEpileptic), !is.na(Gen2AntiPsych))   

melt_bd <- as.data.table(melt_bd)  %>% melt(id.vars = "BrainPAD") %>%
  mutate(moodStabilizer =
                  case_when(
                    variable == "Li" & value == "No" ~ 'No Lithium',
                    variable == "Li" & value == "Yes" ~ 'On Lithium',
                    variable == "AntiEpileptic" & value == "No"~ 'No Antiepileptics',
                    variable == "AntiEpileptic" & value == "Yes" ~ 'On Antiepileptics', 
                    variable == "Gen2AntiPsych" & value == "No"~ 'No SGA',
                    variable == "Gen2AntiPsych" & value == "Yes" ~ 'On SGA', 
                  ))

jpeg(filename='figures/med_usage_on_brainPAD.jpeg', width = 600, height = 600)
melt_bd %>%  mutate(moodStabilizer = factor(moodStabilizer, levels = c("On Lithium", 'No Lithium', 'On AntiEpileptics', 'No AntiEpileptics', "On SGA", 'No SGA'))) %>% 
  ggplot(mapping = aes(x=factor(moodStabilizer), y = BrainPAD , fill=factor(moodStabilizer)))+ ylim(-45,45)+geom_violin( alpha= 0.3 )+ 
  geom_label(data = brainPADs, aes(x = factor(moodStabilizer), y = brainPAD_mean, label =paste(brainPAD_mean)), alpha= 0.2,
  nudge_x = 0,
  nudge_y = 40, size = 6) +   xlab("Medication") + ylab("Brain-PAD (years)") +theme_classic()+ theme(legend.position = "none")+scale_fill_manual(values=c("navy", "lightblue", "red"  ,"pink", "darkgreen", "lightgreen", "grey"))  + 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "red", size = 0.25)  + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18, angle = 90, hjust = 1, vjust = 0.5),
 axis.text.y=element_text(size=18)) +
  geom_vline(xintercept = seq(2.5, 6.5, 2), linetype = "dashed", color = "black", size = 0.5) +  # Add lines between the violins
  theme(
    axis.title.x = element_text(size = 20),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(size = 18)
  ) +scale_x_discrete(labels = function(x) paste(x, "\nN=", brainPADs$count[match(x, levels(factor(brainPADs$moodStabilizer)))], sep = ""))
dev.off()  
```

### Recode Li and AntiEpileptics into four diff levels in a single variable
 

 
```{r}
# One option is to code these as NOT levels, but rather as variables
BD_only <- BD_only %>%
  mutate(moodStabilizer = 
                  case_when(
                    Li == 0 & AntiEpileptic == 0 & Gen2AntiPsych == 0 ~ 'None',
                    Li == 1 & AntiEpileptic == 0 & Gen2AntiPsych == 0 ~ 'Li', # I would have recode this as AntiE = 1 and AntiPsych = 1
                    Li == 0 & AntiEpileptic == 1 & Gen2AntiPsych == 0 ~ 'AED',
                    Li == 0 & AntiEpileptic == 0 & Gen2AntiPsych == 1 ~ 'SGA',
                    Li == 1 & AntiEpileptic == 1 & Gen2AntiPsych == 1 ~ 'AED + SGA + Li',
                    Li == 1 & AntiEpileptic == 1& Gen2AntiPsych == 0 ~ 'AED + Li',
                    Li == 0 & AntiEpileptic == 1 & Gen2AntiPsych == 1 ~ 'AED + SGA',
                    Li == 1 & AntiEpileptic == 0 & Gen2AntiPsych == 1 ~ 'SGA + Li', 
                  ))   
 
med_count<-as.data.table(BD_only) %>% group_by(moodStabilizer) %>% tally() %>% na.omit() #%>% kable() # how many data points in each group?
 
BD_only %>% as.data.table() %>% group_by(moodStabilizer) %>% select(Li, AntiEpileptic, Gen2AntiPsych) %>% as.data.table() %>% melt(id.vars = c('Li', 'Gen2AntiPsych', 'AntiEpileptic'))
```

### Table 2 Count of medication variable with 8 levels
```{r}
# Lithium
li_total_available<-df %>%  filter(Dx == 1)%>% group_by(Li) %>% select(Li) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
med_count%>% mutate(prop = med_count$n/li_total_available$n*100 ) 

# Anti-Epileptic
AntiEpileptic_total_available<- df %>%  filter(Dx == 1)%>% group_by(AntiEpileptic) %>% select(AntiEpileptic) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
med_count%>% mutate(prop = med_count$n/AntiEpileptic_total_available$n*100 ) 

# SGA
Gen2AntiPsych_total_available<- df %>%  filter(Dx == 1)%>% group_by(Gen2AntiPsych) %>% select(Gen2AntiPsych) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
med_count%>% mutate(prop = med_count$n/Gen2AntiPsych_total_available$n*100 ) 
 
```


# Relevel factors and set reference group to none
```{r}
# relevel factors - set ref group to none
BD_only$moodStabilizer<- as.factor(BD_only$moodStabilizer)
BD_only$moodStabilizer <- factor(BD_only$moodStabilizer, levels = c("None", "AED", 'AED + SGA', 'AED + Li',  "Li", "SGA", 'SGA + Li', 'AED + SGA + Li'))

# Check the levels to ensure the desired order
levels(BD_only$moodStabilizer)
```
#

```{r}
m1<-lmer(BrainPAD ~   Age + Sex   +moodStabilizer + (1 | Site), data = BD_only)
summary(m1) 
# 3.59e-05
# chi-square test
```
 

### Est. Marginal Means of antiE, Li, 2ndGen (recoded levels under moodstabilizer factor)
```{r}

emm.s<-emmeans(m1, trt.vs.ctrl ~   moodStabilizer)

counts <- BD_only %>% 
  filter(!is.na(moodStabilizer)) %>%
  group_by(moodStabilizer) %>% 
  summarize(count = n())

jpeg(file="figures/emmeans_medication.jpeg", width = 600, height = 600) 
p <- plot(emm.s,  comparisons = TRUE, , arrow.size = 5) +   
  theme_classic() +
  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18)
  )  + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabilizer)))], sep = "")) + scale_x_continuous("Brain-PAD (years)")+ coord_flip()  + ylab("Medication")
p
 
dev.off()
```

```{r}
# Calculate Bonferroni-adjusted pairwise comparisons
all_pairs <- pairs(emm.s, infer = c(TRUE, TRUE), adjust = "bonferroni")

# Extract results for specific contrast ("AED") with Bonferroni adjustment
contrast_pairs <- pairs(emm.s, contrast = list("AED"), adjust = "bonferroni")

# Convert results to a tibble
pairwise_df <- as_tibble(as.data.frame(all_pairs))

# Round estimate and add significance levels
pairwise_df <- pairwise_df %>%
  mutate(
    estimate = round(estimate, 3),
    significance = case_when(
      p.value < 0.001 ~ '***',
      p.value < 0.01 ~ '**',
      p.value < 0.05 ~ '*',
      TRUE ~ ''
    ))


generate_plot <- function(data, xticks_order) {
  data %>%
    ggplot(aes(x = factor(contrast, levels = rev(xticks_order)), y = estimate)) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0), fill = "lightgrey") +  # Shading below 0
    geom_point(color = 'red', size = 3.5) +
    geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(width = 0.75)) +
    geom_text(aes(label = significance), vjust = -0.3, size = 3) +  # Add this line
    labs(x = "Pairwise Comparisons", y = "Estimated Brain-PAD (years)") +
    ylim(-8, 8) + 
    theme_minimal() +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x = element_text(size = 10, angle = 0, hjust = 1, vjust = 0.5),
      axis.text.y = element_text(size = 10),
      legend.title = element_blank()
    ) +
    coord_flip()
}
```

```{r}
 
replace_ref_contrast <- function(df, med) {
  df %>%
    mutate(
      lower.CL = ifelse(grepl(med, contrast), lower.CL, lower.CL * -1),
      upper.CL = ifelse(grepl(med, contrast), upper.CL, upper.CL * -1),
      estimate = ifelse(grepl(med, contrast), estimate, estimate * -1),
      contrast = ifelse(grepl(med, contrast), contrast , gsub("(.+) - (.+)", "\\2 - \\1", contrast))
    )
}

add_rows_to_df <- function(data_frame, contrast_pattern, specific_contrast) {
  filtered_rows <- replace_ref_contrast(data_frame, contrast_pattern) %>%
    filter(grepl(specific_contrast, contrast))
  return(filtered_rows)
 }

add_rows_to_df_and_rbind <- function(df, col1, col2) {
  r <- add_rows_to_df(df, col1, col2)
  return(r)
}
```

```{r}
columns_to_add <- c(
  "\\(AED \\+ SGA\\)"  , "None - \\(AED \\+ SGA\\)",
  "\\(AED \\+ Li\\)"   , "None - \\(AED \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "None - \\(AED \\+ SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "SGA - \\(AED \\+ SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "Li - \\(AED \\+ SGA \\+ Li\\)"
)

# generates a sequence of numbers starting from 1 to the length of columns_to_add 
seq_cols <- seq(1, length(columns_to_add), by = 2)

# do.call function is used to combine these data frames into a single data frame
none_vs_drug <- do.call(rbind, lapply(seq_cols, function(i) {
  add_rows_to_df_and_rbind(pairwise_df, columns_to_add[i], columns_to_add[i + 1])
}))

jpeg(file="figures/pairwise_emmeans_AED_ref.jpeg", width = 1200, height = 800,  res = 200 ) 
replace_ref_contrast(pairwise_df, "None - AED")%>% filter(str_detect(contrast, "- AED|- \\(AED \\+ SGA\\)|- \\(AED \\+ Li\\)|- \\(AED \\+ SGA \\+ Li\\)")) %>% bind_rows(., none_vs_drug) %>% generate_plot(., c('None - AED', 'Li - AED', 'SGA - AED', '(AED + SGA) - AED' , '(AED + Li) - AED', '(SGA + Li) - AED'	, '(AED + SGA + Li) - AED', 'None - (AED + SGA)', 'Li - (AED + SGA)', 'SGA - (AED + SGA)','(AED + SGA) - (AED + SGA)', '(AED + Li) - (AED + SGA)',  '(SGA + Li) - (AED + SGA)', '(AED + SGA + Li) - (AED + SGA)',  'None - (AED + Li)', 'Li - (AED + Li)', 'SGA - (AED + Li)', '(AED + SGA) - (AED + Li)' , '(SGA + Li) - (AED + Li)', '(AED + SGA + Li) - (AED + Li)', 'None - (AED + SGA + Li)', 'SGA - (AED + SGA + Li)', 'Li - (AED + SGA + Li)')) 
dev.off()
 
```


```{r}
columns_to_add <- c(
"\\(AED \\+ Li\\)", "None - \\(AED \\+ Li\\)",
  "\\(SGA \\+ Li\\)", "None - \\(SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "None - \\(AED \\+ SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "AED - \\(AED \\+ SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "SGA - \\(AED \\+ SGA \\+ Li\\)"
)
# generates a sequence of numbers starting from 1 to the length of columns_to_add 
seq_cols <- seq(1, length(columns_to_add), by = 2)
# do.call function is used to combine these data frames into a single data frame
none_vs_drug <- do.call(rbind, lapply(seq_cols, function(i) {
  add_rows_to_df_and_rbind(pairwise_df, columns_to_add[i], columns_to_add[i + 1])
}))

jpeg(file="figures/pairwise_emmeans_Li_ref.jpeg", width = 1200, height = 800,  res = 200 ) 
replace_ref_contrast(pairwise_df, "- Li") %>% filter(str_detect(contrast, "- Li|- \\(AED \\+ Li\\)|- \\(SGA \\+ Li\\)"))   %>%  bind_rows(., none_vs_drug) %>% generate_plot(., c('None - Li', 'AED - Li', 'SGA - Li', '(AED + SGA) - Li' , '(AED + Li) - Li', '(SGA + Li) - Li'	, '(AED + SGA + Li) - Li', 'None - (AED + Li)', 'Li - (AED + Li)', 'SGA - (AED + Li)','(AED + SGA) - (AED + Li)',    '(SGA + Li) - (AED + Li)', '(AED + SGA + Li) - (AED + Li)',  'None - (SGA + Li)', 'Li - (SGA + Li)', 'SGA - (SGA + Li)', '(AED + SGA) - (SGA + Li)' , '(AED + Li) - (SGA + Li)', '(AED + SGA + Li) - (SGA + Li)', 'None - (AED + SGA + Li)', 'AED - (AED + SGA + Li)', 'SGA - (AED + SGA + Li)')) 
dev.off()
```

```{r}
columns_to_add <- c(
"\\(AED \\+ Li\\)", "None - \\(AED \\+ SGA\\)",
  "\\(SGA \\+ Li\\)", "None - \\(SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "None - \\(AED \\+ SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "AED - \\(AED \\+ SGA \\+ Li\\)",
  "\\(AED \\+ SGA \\+ Li\\)"  , "Li - \\(AED \\+ SGA \\+ Li\\)"
)
# generates a sequence of numbers starting from 1 to the length of columns_to_add 
seq_cols <- seq(1, length(columns_to_add), by = 2)

# do.call function is used to combine these data frames into a single data frame
none_vs_drug <- do.call(rbind, lapply(seq_cols, function(i) {
  
add_rows_to_df_and_rbind(pairwise_df, columns_to_add[i], columns_to_add[i + 1])
}))

jpeg(file="figures/pairwise_emmeans_SGA_ref.jpeg", width = 1200, height = 800,  res = 200 ) 
replace_ref_contrast(pairwise_df, "- SGA")%>% filter(str_detect(contrast, "- SGA|- \\(AED \\+ SGA\\)|- \\(SGA \\+ Li\\)"))  %>%  bind_rows(., none_vs_drug)  %>% generate_plot(., c('None - SGA', 'AED - SGA', 'Li - SGA', '(AED + SGA) - SGA', '(AED + Li) - SGA' ,  '(SGA + Li) - SGA'	, '(AED + SGA + Li) - SGA', 'None - (AED + SGA)', 'Li - (AED + SGA)', '(AED + Li) - (AED + SGA)',  '(SGA + Li) - (AED + SGA)','(AED + SGA + Li) - (AED + SGA)',   'None - (SGA + Li)', 'AED - (SGA + Li)', 'Li - (SGA + Li)', 'SGA - (SGA + Li)',    '(AED + SGA + Li) - (SGA + Li)', 'None - (AED + SGA + Li)', 'AED - (AED + SGA + Li)', 'Li - (AED + SGA + Li)')) 
dev.off() 
```


### Specify estimated marginal means reference group = Li
```{r}
emms.li.ref <- emmeans(m1, trt.vs.ctrl ~ moodStabilizer, ref = 6)
#emms.li.ref$contrasts %>% kable(digits = 3) 
```

# collapsed moodstabilizer LMM (AED & Li+AED  & AED+SGA) 
```{r}
# table (any lithium?)
BD_only$moodStabCollapseAntiE<- fct_collapse(BD_only$moodStabilizer ,AED = c( "AED + SGA"  , "Li + AED", "AED"))
#BD_only %>% group_by(moodStabCollapse) %>% tally() %>% kable() # how many data points in each group?

m.collapse.aed<-lmer(BrainPAD ~   Age + Sex   +moodStabCollapseAntiE + (1 | Site), data = BD_only)
summary(m.collapse.aed)
```
```{r}
emm.s.aed<-emmeans(m.collapse.aed, trt.vs.ctrl ~   moodStabCollapseAntiE)
# calculate -lvalues and CI for pairwise comparisons 
# adjust the p-values for multiple comparisons 
pairs(emm.s, infer = c(TRUE, TRUE), adjust = "bonferroni")  

# Get the counts for each medication group
counts <- BD_only %>% 
  filter(!is.na(moodStabCollapseAntiE)) %>%
  group_by(moodStabCollapseAntiE) %>% 
  summarize(count = n())

#jpeg(file="figures/emmeans_medication_collapse.jpeg")
plot(emm.s.aed,ylab="Medication",xlab="Brain-PAD", comparisons = TRUE, horizontal=F)+ theme_classic()  + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18),
 axis.text.y=element_text(size=18))    + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabCollapseAntiE)))], sep = ""))  + 
  scale_x_continuous("Brain-PAD (years)")+ coord_flip()  + ylab("Medication")
# dev.off() 

```

# collapsed moodstabilizer LMM (Li & Li+AntiE  & Li+AntiPsych) 

```{r}
# table (any lithium?)
BD_only$moodStabCollapse <- fct_collapse(BD_only$moodStabilizer ,Li = c( "Li + SGA"  , "Li + AED", "Li"))
#BD_only %>% group_by(moodStabCollapse) %>% tally() %>% kable() # how many data points in each group?

m.collapse.li<-lmer(BrainPAD ~   Age + Sex   +moodStabCollapse + (1 | Site), data = BD_only)
summary(m.collapse.li)
```



### Specify estimated marginal means reference group = none for collapsed moodstabilizer LMM 
```{r}

# Get the counts for each medication group
counts <- BD_only %>% 
  filter(!is.na(moodStabCollapse)) %>%
  group_by(moodStabCollapse) %>% 
  summarize(count = n())

m.collapse.li<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse )  

#jpeg(file="figures/emmeans_medication_collapse.jpeg")
plot(m.collapse.li,ylab="Medication",xlab="Brain-PAD", comparisons = TRUE, horizontal=F)+ theme_classic()  + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18),
 axis.text.y=element_text(size=18))    + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabCollapse)))], sep = ""))  + 
  scale_x_continuous("Brain-PAD (years)")+ coord_flip()  + ylab("Medication")
# dev.off() 
# m.collapse.li.ref$contrasts  %>% kable(digits = 3)
# m.collapse.li.ref$contrasts 
```


### Specify estimated marginal means reference group = Li for collapsed moodstabilizer LMM 
```{r}
m.collapse.li<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse ) 
m.collapse.li.ref<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse, ref = 6 ) 

jpeg(file="figures/emmeans_medication_any_li.jpeg", width = 600, height = 600) 
p  <-plot(m.collapse.li.ref, comparisons = TRUE) + 
  theme_classic() + 
  scale_x_continuous("Brain-PAD (years)") + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabCollapse)))], sep = ""))  + ylab("Medication") + coord_flip() +  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18) 
  ) 
# Add the text labels outside of the bars
p #+ geom_text_repel(data = counts, aes(x = Inf, y = as.character(moodStabCollapse), label = count), color = "black", size = 6,  , hjust = 0, nudge_x = 0.05)
dev.off()

m.collapse.li.ref$contrasts  
```

### LMM and Estimated Marginal Means Li*Antiepiletpics*2ndGenAntiPsych (Not recoded - each med is separate term in model)
This model has 3 terms AntiEpileptic, Li, and Gen2AntiPsych and their pairwise interaction terms.  

The factors with levels to compare are on the right hand side (AntiEpileptics and Gen2AntiPsych). 
- returns emmeans results on the model scale (e.g., estimated mean brainPAD for each AntiEpileptic*Gen2AntiPsych combination, which is a total of 4)

```{r}
m.interaction <-lmer(BrainPAD ~   Age + Sex   + AntiEpileptic*Li*Gen2AntiPsych+ (1 | Site), data = BD_only)
summary(m.interaction) 

```

### Visualize interactions 
Being on AntiE does not always corr with higher brainAge
Those on AntiE have higher brainPAD when NOT on Li compared to those on AntiPsych when NOT on Li
Those on AntiPsych have higher brainPAD when on Li comapred to those on AntiE when on Li
 
```{r}
# formula needs x var on right hand side, and trace factor on the left
# trace factor is the var(s) used to define the lines/curves
# NOT an interaction plot of RAW data, they are plots of model predictions
# emmip(m.interaction,  AntiEpileptic  *Gen2AntiPsych ~  Li , CIs = TRUE) + scale_colour_discrete(name="Medication", labels=c("Neither","On AntiEpileptic", "On Gen2AntiPsych", "Both" )) +theme_classic()   + scale_x_continuous(breaks = seq(0, 1, len = 2)) 
# emmeans(m.interaction,  AntiEpileptic  *Gen2AntiPsych ~  Li , CIs = TRUE) 
# https://aosmith.rbind.io/2019/03/25/getting-started-with-emmeans/
# https://cran.r-project.org/web/packages/emmeans/vignettes/interactions.html#simple
```

### Compare levels of Li (1 vs 0) within each level of AntiE (1 vs 0) and Gen2Anti (1 vs 0) in combination  
This is an extension of what we see in the LMM above. where interaction between Li*AntiE is significant
```{r}
# comparison of On Li vs. Not on Li by Gen2AntiPsych
#emmeans(m1,   pairwise ~ Li |Gen2AntiPsych)  

# comparison of On Li vs. Not on Li by AntiE
# emmeans(m1,   pairwise ~  Li | AntiEpileptic)  

# comparison of On Li vs. Not on Li by the main effect of the combination of each level of AntiE, Gen2 (1 vs 0)
# factors with levels to compare are on the right hand side.  
emms.interaction <- emmeans(m.interaction,   pairwise ~  Li | AntiEpileptic:Gen2AntiPsych)
emms.interaction$contrasts  
pairs(emms.interaction, reverse = T) %>% as.data.frame() %>% mutate_if(is.numeric, funs(as.character(signif(., 3)))) # %>% kable(digits =3)
# confint(emms.s, level = 0.90, adjust = "scheffe")
 
```
 



```{r}
 
# as long as AntiPsych = 1 (see second row), we contrast taking vs. not taking Li
# we combine the effect of antiE and antiPsych

# antipsych = 1, antiE = 1, Li = 1 vs.   antipsych = 1, antiE = 1, Li = 0 (A+B+Li vs. A+B) this is different from (A+B vs.Li+A / A+B vs. Li+B)
# antipsych = 1, antiE = 0, Li = 1 vs.   antipsych = 1, antiE = 0, Li = 0
m1<-lmer(BrainPAD ~   Age + Sex   + AntiEpileptic*Li*Gen2AntiPsych+ (1 | Site), data = BD_only)

emms.s <- emmeans(m1,   pairwise ~ Li | AntiEpileptic+Gen2AntiPsych )
pairs(emms.s, reverse = T)
emmip(m1, ~ Li | Gen2AntiPsych, CIs = TRUE)
#mutate(level2 = case_when(Domain == 'Current pharmacological treatment' & level3 == 'Total psychiatric medications (count)' ~ "Current Drug Class",
```

```{r}
as.data.frame(emm.s$contrasts)  %>% write_xlsx(., 'supp_tables/emmeans_mood_stablizer.xlsx')
```
 


```{r}
# Relationship of Li to BrainPAD adjusting for covariates, excluding users of AntiEpileptics
# Li, b= -3.25, t= -3.85, p < 0.001
excludeNonAntiEpileptic_df <- BD_only %>%
  filter(AntiEpileptic == 1) #only AntiEpileptic  users
m1<-lmer(BrainPAD ~   Age + Sex   + Li + (1 | Site), data = excludeNonAntiEpileptic_df)
summary(m1) 
```

```{r}
# Relationship of Li to BrainPAD adjusting for covariates
# b=-2.69, t=-5.35, p<0.001 
m1 <-lmer(BrainPAD ~   Age + Sex   + Li + (1 | Site), data = BD_only)
summary(m1)
```
 

```{r}
# Relationship of AntiEpileptic to BrainPAD adjusting for covariates
# b=2.31, t=4.71, p<0.001
 
m1 <-lmer(BrainPAD ~   Age + Sex   + AntiEpileptic + (1 | Site), data = BD_only)
summary(m1)
```


```{r}

# Relationship of AntiEpileptic to BrainPAD adjusting for covariates, excluding non users of Li
# AntiEpileptic, b=1.04, t=1.22, p = NS
excludeNonLi_df <- df %>%
  filter(Li == 1) #only Lithium  users 
m1 <-lmer(BrainPAD ~   Age + Sex   + AntiEpileptic + (1 | Site), data = excludeNonLi_df)
summary(m1) 
```


```{r}
# Interaction between Li and Anti-Epileptic to BrainPAD adjusting for covariates
# interaction term, p= NS
m1 <-lmer(BrainPAD ~   Age + Sex    +(Li*AntiEpileptic) + (1 | Site), data = BD_only)
summary(m1) 
```
