---
title: "R Notebook"
output: html_notebook
---
```{r, message = FALSE}
library("dplyr")
library("tidyverse")
library("DescTools")
library("ggplot2")
library("readxl")
library("lme4")
library("ModelMetrics")
library("merTools")
library("lmtest")
library("emmeans")
library("haven")
library("lmerTest")
library("metaforest")
library("rstatix")
library("ech")
library("labelled")
 
libs <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest", "metaforest", "rstatix",  'ech', 'labelled')
#require(libs)
lapply(libs, require, character.only = TRUE)
 
# _fixed dataset has removed any misalignment of subjects across sites
path = file.path( "Master_BD_080720_withPredictions_withNewHajek_fixed.sav")
df = read_sav(path)

df <- df %>% 
  mutate_if(is.character, list(~na_if(.,""))) %>%
  filter(!SubjID %in%  c('con_11',  'pat_13')) %>%
    arrange(Site) 
siteLoc <- read_excel("DescriptiveData_080720_2.xlsx", sheet=3)

# create function to replace spaces with underscores
replaceSpace <- function(dataframe) {
   gsub(x = names(dataframe), pattern = " ", replacement = "_") 
} 

names(siteLoc) <- replaceSpace(siteLoc)
names(df) <- replaceSpace(df) 

# merge the main dataset with the Site Location
df <- merge(df, siteLoc, by=c("PIName","Site"  )) 
df <- df %>% 
  filter(Age <= 75  & Age  >=18) 
BD_only <- df %>% filter(Dx==1) 

length(colnames(BD_only))-23

```


### Plot BrainPAD Lithium, Antiepileptics, SGA vs. not on each of those 

```{r}
meanBrainPAD_Li <- df %>% 
  filter(!is.na(Li)) %>%
  group_by(Li) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2)) %>%
  `colnames<-`(c("moodStabilizer", "brainPAD_mean"))   %>%
  mutate(moodStabilizer =
                  case_when(
                    moodStabilizer == 0   ~ 'No Lithium',
                    moodStabilizer == 1  ~ 'On Lithium', 
                  ))

meanBrainPAD_AntiE <- df %>% 
  filter( !is.na(AntiEpileptic)) %>%
  group_by(AntiEpileptic) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2))  %>%
  `colnames<-`(c("moodStabilizer", "brainPAD_mean"))%>%
  mutate(moodStabilizer =
                  case_when(
                    moodStabilizer == 0   ~ 'No AntiEpileptics',
                    moodStabilizer == 1  ~ 'On AntiEpileptics', 
                  )) 
 
meanBrainPAD_SGA <- df %>% 
  filter( !is.na(Gen2AntiPsych)) %>%
  group_by(Gen2AntiPsych) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2))  %>%
  `colnames<-`(c("moodStabilizer", "brainPAD_mean"))%>%
  mutate(moodStabilizer =
                  case_when(
                    moodStabilizer == 0   ~ 'No SGA',
                    moodStabilizer == 1  ~ 'On SGA', 
                  )) 
brainPADs <-  rbind(meanBrainPAD_Li, meanBrainPAD_AntiE, meanBrainPAD_SGA)

melt_bd <- unlabelled(BD_only) %>% dplyr::select(BrainPAD, Li, AntiEpileptic, Gen2AntiPsych) %>%
  filter(!is.na(Li), !is.na(AntiEpileptic), !is.na(Gen2AntiPsych))   

melt_bd <- melt_bd  %>% melt(id.vars = "BrainPAD") %>%
  mutate(moodStabilizer =
                  case_when(
                    variable == "Li" & value == "No" ~ 'No Lithium',
                    variable == "Li" & value == "Yes" ~ 'On Lithium',
                    variable == "AntiEpileptic" & value == "No"~ 'No AntiEpileptics',
                    variable == "AntiEpileptic" & value == "Yes" ~ 'On AntiEpileptics', 
                    variable == "Gen2AntiPsych" & value == "No"~ 'No SGA',
                    variable == "Gen2AntiPsych" & value == "Yes" ~ 'On SGA', 
                  ))

jpeg(filename='figures/med_usage_on_brainPAD.jpeg', width = 600, height = 600)
melt_bd %>%  mutate(moodStabilizer = factor(moodStabilizer, levels = c("On Lithium", 'No Lithium', 'On AntiEpileptics', 'No AntiEpileptics', "On SGA", 'No SGA'))) %>% 
  ggplot(mapping = aes(x=factor(moodStabilizer), y = BrainPAD , fill=factor(moodStabilizer)))+ ylim(-45,45)+geom_violin( alpha= 0.3 )+ 
  geom_label(data = brainPADs, aes(x = factor(moodStabilizer), y = brainPAD_mean, label = paste(  brainPAD_mean)), alpha= 0.2,
  nudge_x = 0,
  nudge_y = 40, size = 8) +   xlab("Medication") +theme_classic()+ theme(legend.position = "none")+scale_fill_manual(values=c("navy", "lightblue", "red"  ,"pink", "darkgreen", "lightgreen", "grey"))  + 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "red", size = 0.25)  + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18, angle = 90, hjust = 1, vjust = 0.5),
 axis.text.y=element_text(size=18)) 
dev.off()

# Only lithium yes vs. no  
# jpeg(filename='figures/lithium_usage_on_brainPAD.jpeg')
# BD_only %>%
#   filter(!is.na(Li), !is.na(AntiEpileptic) ) %>%
#   ggplot(mapping = aes(x=factor(Li, labels = c('No','Yes')), y = BrainPAD , fill=factor(Li)))+ ylim(-45,45)+geom_violin(, alpha= 0.2)+ 
#   geom_label(data = meanBrainPAD_Li, aes(x = factor(moodStabilizer, labels = c('No','Yes')), y = brainPAD_mean, label = paste(  brainPAD_mean)), alpha= 0.2) +   xlab("Lithium Usage") +theme_classic()+ theme(legend.position = "none")+scale_fill_manual(values=c("darkgray","red" ))  
# dev.off()
```

### Recode Li and AntiEpileptics into four diff levels in a single variable
```{r}
# BD_only <- BD_only %>%
#   mutate(moodStabilizer = 
#                   case_when(
#                     Li == 0 & AntiEpileptic == 0 ~ 'none',
#                     Li == 0 & AntiEpileptic == 1 ~ 'AntiE',
#                     Li == 1 & AntiEpileptic == 0 ~ 'Li',
#                     Li == 1 & AntiEpileptic == 1 ~ 'both',
#                     
#                   ))  
 
# relevel factors - set ref group to none
# BD_only$moodStabilizer<- as.factor(BD_only$moodStabilizer)
# BD_only$moodStabilizer <- relevel(BD_only$moodStabilizer,"none")
# levels(BD_only$moodStabilizer)
#BD_only %>% group_by(moodStabilizer) %>% tally() %>% write_xlsx(., "supp_tables/mood_stabilizer_count.xlsx")
```

 
```{r}
# One option is to code these as NOT levels, but rather as variables
BD_only <- BD_only %>%
  mutate(moodStabilizer = 
                  case_when(
                    Li == 0 & AntiEpileptic == 0 & Gen2AntiPsych == 0 ~ 'none',
                    Li == 1 & AntiEpileptic == 0 & Gen2AntiPsych == 0 ~ 'Li', # I would have recode this as AntiE = 1 and AntiPsych = 1
                    Li == 0 & AntiEpileptic == 1 & Gen2AntiPsych == 0 ~ 'AntiEpileptic',
                    Li == 0 & AntiEpileptic == 0 & Gen2AntiPsych == 1 ~ 'Gen2AntiPsych',
                    Li == 1 & AntiEpileptic == 1 & Gen2AntiPsych == 1 ~ 'all',
                    Li == 1 & AntiEpileptic == 1& Gen2AntiPsych == 0 ~ 'Li&AntiEpileptic',
                    Li == 0 & AntiEpileptic == 1 & Gen2AntiPsych == 1 ~ 'AntiEpileptic&Gen2AntiPsych',
                    Li == 1 & AntiEpileptic == 0 & Gen2AntiPsych == 1 ~ 'Li&Gen2AntiPsych', 
                  ))   
 
med_count<-BD_only %>% group_by(moodStabilizer) %>% tally() %>% na.omit() #%>% kable() # how many data points in each group?
```

### Table 2 Count of medication variable with 8 levels
```{r}
# Lithium
li_total_available<-df %>%  filter(Dx == 1)%>% group_by(Li) %>% select(Li) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
med_count%>% mutate(prop = med_count$n/li_total_available$n*100 ) 

# Anti-Epileptic
AntiEpileptic_total_available<- df %>%  filter(Dx == 1)%>% group_by(AntiEpileptic) %>% select(AntiEpileptic) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
med_count%>% mutate(prop = med_count$n/AntiEpileptic_total_available$n*100 ) 


# SGA
Gen2AntiPsych_total_available<- df %>%  filter(Dx == 1)%>% group_by(Gen2AntiPsych) %>% select(Gen2AntiPsych) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
med_count%>% mutate(prop = med_count$n/Gen2AntiPsych_total_available$n*100 ) 


```


```{r}
# relevel factors - set ref group to none
BD_only$moodStabilizer<- as.factor(BD_only$moodStabilizer)
BD_only$moodStabilizer <- relevel(BD_only$moodStabilizer,"none")
levels(BD_only$moodStabilizer)
#BD_only %>% group_by(moodStabilizer) %>% tally() %>% write_xlsx(., "supp_tables/mood_stabilizer_count.xlsx")
```


```{r}
m1<-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 +moodStabilizer + (1 | Site), data = BD_only)
summary(m1) 
# 3.59e-05
```
 

### Est. Marginal Means of antiE, Li, 2ndGen (recoded levels under moodstabilizer factor)
```{r}
# pairwise comparisons between the groups in this model
emm.s<-emmeans(m1, trt.vs.ctrl ~ moodStabilizer)
jpeg(file="figures/emmeans_medication.jpeg", width = 600, height = 600)
# We have 
#  AntiE&2ndGenAntiPsyc - Li                   5.051 0.922 1316   5.478  <.0001
#  AntiE&2ndGenAntiPsyc - Li&2ndGenAntiPsyc    3.335 1.048 1307   3.182  0.0322
#  AntiE&2ndGenAntiPsyc - Li&AntiE             4.539 1.098 1313   4.133  0.0010
plot(emm.s, comparisons = TRUE) + 
  theme_classic() + 
  scale_x_continuous("Brain-PAD") + 
 # ggtitle("Mean estimate across medication and 95% CI of BrainPAD") + 
  scale_y_discrete(name = "Medication", labels= c("None", "All", "AntiEpileptic", "AntiEpileptics + SGA", "SGA", "Li", "Li + AntiEpilpetic", "Li + SGA")) + coord_flip() + 
  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18)
  )
dev.off()
```

```{r}
#emms.li.ref$contrasts %>% kable(digits = 3)
```


### Specify estimated marginal means reference group = Li
```{r}
emms.li.ref <- emmeans(m1, trt.vs.ctrl ~ moodStabilizer, ref = 6)
emms.li.ref$contrasts %>% kable(digits = 3)

```
# collapsed moodstabilizer LMM (Li & Li+AntiE  & Li+AntiPsych) 

```{r}
# table (any lithium?)
BD_only$moodStabCollapse <- fct_collapse(BD_only$moodStabilizer ,
  Li = c( "Li&Gen2AntiPsych"  , "Li&AntiEpileptic", "Li"))
#BD_only %>% group_by(moodStabCollapse) %>% tally() %>% kable() # how many data points in each group?


m.collapse.li<-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 +moodStabCollapse + (1 | Site), data = BD_only)
summary(m.collapse.li)
```



### Specify estimated marginal means reference group = none for collapsed moodstabilizer LMM 
```{r}
m.collapse.li<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse ) 
#m.collapse.li.ref<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse) 

jpeg(file="figures/emmeans_medication_collapse.jpeg")
plot(m.collapse.li,ylab="Medication",xlab="Brain-PAD", comparisons = TRUE, horizontal=F)+ theme_classic()  + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18),
 axis.text.y=element_text(size=18))    # +ggtitle("Mean estimate across medication and 95% CI of BrainPAD") 
dev.off()

m.collapse.li.ref$contrasts  %>% kable(digits = 3)
```


### Specify estimated marginal means reference group = Li for collapsed moodstabilizer LMM 
```{r}
m.collapse.li<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse ) 
m.collapse.li.ref<- emmeans(m.collapse.li, trt.vs.ctrl ~ moodStabCollapse, ref = 6 ) 

jpeg(file="figures/emmeans_medication_any_li.jpeg", width = 600, height = 600)

 
# plot(m.collapse.li.ref,  xlab="Brain-PAD", comparisons = TRUE, horizontal=F)+ggtitle("Mean estimate across medication and 95% CI of BrainPAD") + theme_classic() + scale_x_discrete(name = "Medication", labels= c("None", "All", "AntiEpileptic", "AntiEpileptics + SGA", "SGA", "Li"))  + 
#  
#    coord_flip() + 
#   theme(
#     axis.title.x = element_text(size = 20), 
#     plot.title = element_text(size = 20, face = "bold"), 
#     axis.title.y = element_text(size = 20), 
#     axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
#     axis.text.y = element_text(size = 18)
#   )


plot(m.collapse.li.ref, comparisons = TRUE) + 
  theme_classic() + 
  scale_x_continuous("Brain-PAD") + 
 # ggtitle("Mean estimate across medication and 95% CI of BrainPAD") + 
  scale_y_discrete(name = "Medication", labels = c("None", "All", "AntiEpileptic", "AntiEpileptics + SGA", "SGA", "Li" )) + 
  coord_flip() + 
  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18)
  )

dev.off()

m.collapse.li.ref$contrasts  %>% kable(digits = 3)
```

### LMM and Estimated Marginal Means Li*Antiepiletpics*2ndGenAntiPsych (Not recoded - each med is separate term in model)
This model has 3 terms AntiEpileptic, Li, and Gen2AntiPsych and their pairwise interaction terms.  

The factors with levels to compare are on the right hand side (AntiEpileptics and Gen2AntiPsych). 
- returns emmeans results on the model scale (e.g., estimated mean brainPAD for each AntiEpileptic*Gen2AntiPsych combination, which is a total of 4)

```{r}
m.interaction <-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + AntiEpileptic*Li*Gen2AntiPsych+ (1 | Site), data = BD_only)
summary(m.interaction)


```

### Visualize interactions 
Being on AntiE does not always corr with higher brainAge
Those on AntiE have higher brainPAD when NOT on Li compared to those on AntiPsych when NOT on Li
Those on AntiPsych have higher brainPAD when on Li comapred to those on AntiE when on Li
 
```{r}
# formula needs x var on right hand side, and trace factor on the left
# trace factor is the var(s) used to define the lines/curves
# NOT aninteraction plot of RAW data, they are plots of model predictions
emmip(m.interaction,  AntiEpileptic  *Gen2AntiPsych ~  Li , CIs = TRUE) + scale_colour_discrete(name="Medication", labels=c("Neither","On AntiEpileptic", "On Gen2AntiPsych", "Both" )) +theme_classic()   + scale_x_continuous(breaks = seq(0, 1, len = 2)) 
emmeans(m.interaction,  AntiEpileptic  *Gen2AntiPsych ~  Li , CIs = TRUE) 
# https://aosmith.rbind.io/2019/03/25/getting-started-with-emmeans/
# https://cran.r-project.org/web/packages/emmeans/vignettes/interactions.html#simple
```

### Compare levels of Li (1 vs 0) within each level of AntiE (1 vs 0) and Gen2Anti (1 vs 0) in combination  
This is an extension of what we see in the LMM above. where interaction between Li*AntiE is significant
```{r}
# comparison of On Li vs. Not on Li by Gen2AntiPsych
#emmeans(m1,   pairwise ~ Li |Gen2AntiPsych)  

# comparison of On Li vs. Not on Li by AntiE
# emmeans(m1,   pairwise ~  Li | AntiEpileptic)  

# comparison of On Li vs. Not on Li by the main effect of the combination of each level of AntiE, Gen2 (1 vs 0)
# factors with levels to compare are on the right hand side.  
emms.interaction <- emmeans(m.interaction,   pairwise ~  Li | AntiEpileptic:Gen2AntiPsych)
emms.interaction$contrasts  
pairs(emms.interaction, reverse = T) %>% as.data.frame() %>% mutate_if(is.numeric, funs(as.character(signif(., 3))))  %>% kable(digits =3)
# confint(emms.s, level = 0.90, adjust = "scheffe")
 
```
# Pairwise contrast  AntiEpileptic:Li | Gen2AntiPsych
```{r}
# as i understand it, we want to collapse these three results into one
# AntiE = 1 vs Li = 1 ->  AntiEpileptic1 Li0 - AntiEpileptic0 Li1           4.465 0.836 1310   5.338  <.0001 (Gen2AntiPsych = 0)
# AntiE = 1 vs Li&AntiE = 1 -> AntiEpileptic1 Li0 - AntiEpileptic1 Li1      3.953 1.029 1308   3.843  0.0007 (Gen2AntiPsych = 0)
# AntiE = 1 vs Li&AntiPsych = 1 ->  AntiEpileptic1 Li0 - AntiEpileptic0 Li1 3.335 1.048 1307   3.182  0.0081 (Gen2AntiPsych = 1)
emms.s <- emmeans(m.interaction,   pairwise ~ AntiEpileptic:Li | Gen2AntiPsych)
emms.s$contrasts 
```
### Pairwise contrast   Li | AntiEpileptic
```{r}
# this combines/collapses the following, in the second row:
# as long as AntiE = 1 (see second row), we contrast taking vs. not taking Li
# we combine the effect of antiE and antiPsych

# antipsych = 0, antiE = 1 (Li + antiE vs. antiE)
# antipsych = 1, antiE = 1 (Li + antiE vs. antiE + antipsych)
 
emms.s <- emmeans(m.interaction,  "Li", counterfact = "AntiEpileptic")
ref_grid(m.interaction)

```



```{r}
 
# as long as AntiPsych = 1 (see second row), we contrast taking vs. not taking Li
# we combine the effect of antiE and antiPsych

# antipsych = 1, antiE = 1, Li = 1 vs.   antipsych = 1, antiE = 1, Li = 0 (A+B+Li vs. A+B) this is different from (A+B vs.Li+A / A+B vs. Li+B)
# antipsych = 1, antiE = 0, Li = 1 vs.   antipsych = 1, antiE = 0, Li = 0
m1<-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + AntiEpileptic*Li*Gen2AntiPsych+ (1 | Site), data = BD_only)

emms.s <- emmeans(m1,   pairwise ~ Li | AntiEpileptic+Gen2AntiPsych )
pairs(emms.s, reverse = T)
emmip(m1, ~ Li | Gen2AntiPsych, CIs = TRUE)
#mutate(level2 = case_when(Domain == 'Current pharmacological treatment' & level3 == 'Total psychiatric medications (count)' ~ "Current Drug Class",
```

```{r}
as.data.frame(emm.s$contrasts) %>%
  format_table() %>% write_xlsx(., 'supp_tables/emmeans_mood_stablizer.xlsx')
```
 


```{r}
# Relationship of Li to BrainPAD adjusting for covariates, excluding users of AntiEpileptics
# Li, b= -3.25, t= -3.85, p < 0.001
excludeNonAntiEpileptic_df <- BD_only %>%
  filter(AntiEpileptic == 1) #only AntiEpileptic  users
m1<-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + Li + (1 | Site), data = excludeNonAntiEpileptic_df)
summary(m1) 
```

```{r}
# Relationship of Li to BrainPAD adjusting for covariates
# b=-2.69, t=-5.35, p<0.001 
m1 <-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + Li + (1 | Site), data = BD_only)
summary(m1)
```
 





```{r}
# Relationship of AntiEpileptic to BrainPAD adjusting for covariates
# b=2.31, t=4.71, p<0.001
 
m1 <-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + AntiEpileptic + (1 | Site), data = BD_only)
summary(m1)
```


```{r}

# Relationship of AntiEpileptic to BrainPAD adjusting for covariates, excluding non users of Li
# AntiEpileptic, b=1.04, t=1.22, p = NS
excludeNonLi_df <- df %>%
  filter(Li == 1) #only Lithium  users 
m1 <-lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + AntiEpileptic + (1 | Site), data = excludeNonLi_df)
summary(m1) 
```


```{r}
# Interaction between Li and Anti-Epileptic to BrainPAD adjusting for covariates
# interaction term, p= NS
m1 <-lmer(BrainPAD ~   Age + Sex + Age2 + Age3  +(Li*AntiEpileptic) + (1 | Site), data = BD_only)
summary(m1) 
```
