---
title: "BD_Generalization"
author: "HXNG"
date: "4/13/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
```{r} 
libraries <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest", "metaforest", "rstatix")
require(libraries)
lapply(libraries, require, character.only = TRUE)
path = file.path( "Master_BD_080720_withPredictions_withNewHajek_fixed.sav")
df = read_sav(path)

# Helper function to rename df columns
replaceSpace <- function(dataframe) {
  gsub(x = names(dataframe), pattern = " ", replacement = "_") 
}

df <- df %>% 
  mutate_if(is.character, list(~na_if(.,""))) %>%
  filter(!SubjID %in%  c('con_11',  'pat_13')) %>% # drop 2 subjects from Haarman
  arrange(Site) # arrange by Site number

siteLoc <- read_excel("DescriptiveData_080720_2.xlsx", sheet=3)

names(siteLoc) <- replaceSpace(siteLoc)
names(df) <- replaceSpace(df) 

df <- merge(df, siteLoc, by=c("PIName","Site"  )) # merge 
df <- df %>% 
  filter(Age <= 75  & Age  >=18)  
BD_only <- df %>% filter(Dx==1) 
```


### Count num people not on any medications
```{r}
site_no_med <-BD_only %>% 
  filter(AntiEpileptic==0, 
         Li==0,AntiDep==0, 
         Gen1AntiPsych==0, 
         Gen2AntiPsych==0) %>%  
  group_by(Site_Locations) %>% tally() 
# site_no_med %>% summarize(n = sum(n))
```

```{r}
site_all_bd <- BD_only %>% dplyr::count(Site_Locations, PIName)
site_all_bd <- merge(site_all_bd, site_no_med, by=c("Site_Locations"))
names(site_all_bd) <- c('Site','PIName','total_BD', 'no_med_BD')
write.csv(site_all_bd,'enigmaBrainAGE_no_med_BD_2022may6.csv',row.names = FALSE)
```

  

```{r}
# Lithium
df %>%  filter(Dx == 1)%>% group_by(Li) %>% select(Li) %>% tally() %>% na.omit() %>% summarize(n = sum(n))

# Anti-Epileptic
df %>%  filter(Dx == 1)%>% group_by(AntiEpileptic) %>% select(AntiEpileptic) %>% tally() %>% na.omit() %>% summarize(n = sum(n))

# SGA
df %>%  filter(Dx == 1)%>% group_by(Gen2AntiPsych) %>% select(Gen2AntiPsych) %>% tally() %>% na.omit() %>% summarize(n = sum(n))
```
 
```{r}
count <- df %>% filter(Dx == 1) %>%  group_by(BDI) %>% select(BDI) %>% tally() %>% na.omit()
count
sum(count$n)
count %>% prop.table() %>% round(3)*100


df %>% 
  group_by(Sex, Dx)%>% 
  dplyr::summarize(MAE = mean(abs(age_prediction - Age)), sdMAE= sd(abs(age_prediction - Age)))  
```


## MAE of Each Site, Separate plot for each Dx*Sex groups
```{r, fig.width=11 }
# Sort by Site Number
# Show separate table with Site Names
AEeachSite <- df %>%
  arrange(Site)%>% 
  group_by(Site, Site_Locations, Sex, Dx,Age)%>% 
  dplyr::summarize(AE = abs(age_prediction - Age)) %>% arrange(Site_Locations) 

for (dx in c(0, 1)) {
  for (sex in c(1,2)){
    if (dx == 0) { 
      dxTitle = "HC"} else{
        dxTitle = "BD"
      }
    if (sex == 1) { 
      sexTitle = "Men"} else{
        sexTitle = "Women"
      }
    
    
    # MAE of Model of HC Men
p<-AEeachSite %>%
  filter(Sex==sex, Dx==dx)%>%
  ggplot(aes(x = factor(Site_Locations) , y=AE))+
  geom_violin() + theme_classic() + stat_summary(fun=mean, geom="point",  color="red", size=2) +xlab('Site')+ggtitle(paste0('Absolute Errors across Sites in ',dxTitle, ' ', sexTitle)) +theme(plot.title = element_text(hjust = 0.51))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
jpeg(filename = paste0('figures/',dxTitle,'_' ,sexTitle,'MAE_by_site.jpeg'))
print(p)
dev.off()

  }
  }
 
```


 

### Distribution of BrainPADs in HC vs. BD
```{r}
jpeg(filename='figures/compare_dist_brainPAD_HC_BD.jpeg', width = 600, height = 600)
df %>%
  dplyr::select(BrainPAD, Dx) %>%
  ggplot(aes(BrainPAD)) +
  geom_density(aes(fill=factor(Dx)), alpha=.25)  + scale_x_continuous("Brain-PAD") + scale_y_continuous("Density")  + 
  scale_fill_manual(name='Group', labels = c("Control", "BD"), 
                    values = c("darkgray","red" )) +    guides(fill=guide_legend(title="Group")) + theme_classic()  +theme(legend.position="bottom") + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18),
 axis.text.y=element_text(size=18)) 
dev.off()
# Average Brain-PADs were −0.80 (SD 9.38) in the control and +0.52 years (SD 9.87) in the BD group.
df %>%
  dplyr::select(BrainPAD, Dx) %>% group_by(Dx) %>%  summarise_at(vars(BrainPAD), list(meanBrainPAD = mean, sd = sd)) %>% round(2)
```
 
### Plot EMMs for HC vs. Dx groups
```{r}

mod <- lmer(BrainPAD ~ Dx + Sex +   Age + Age2 + Age3 + (1 | Site)
           , data = df)    
emm.s <- emmeans(mod, specs = pairwise ~ Dx) 
Means = as.data.frame(emm.s$emmeans)
# The EMM of BrainPADs is -0.95 (SE 0.68, CI: -2.34 - 0.44), for healthy controls and +1.42  years (SE 0.66,  CI: 0.05 - 2.78) for BD individuals.  

mod_fixed <- lm(BrainPAD ~ Dx + Sex +  Age + Age2 + Age3 + Site
           , data = df)   
# Use anova to get f values 
round(as.data.frame(anova(mod_fixed)),3) 
round(EtaSq(mod_fixed), 3) # eta squared  
```

```{r}
df %>% 
  ggplot(mapping = aes(x=factor(Dx, labels = c('HC','BD')), y = BrainPAD, fill=factor(Dx, labels = c('Control','BD')))) + geom_violin(alpha=0.25)  + guides(fill=guide_legend(title="Sex")) + xlab('Sex') + ylim(-60,40)  +scale_fill_manual(values=c("darkgray","red" )) + theme_classic()   +
  geom_pointrange(
    data = Means,
    aes(y = emmean, ymin = lower.CL , ymax = upper.CL),
    size = 0.2,
    color = "brown") #+ geom_line(data = Means, aes(y = emmean, group = 0), size = 0.2, color = 'brown')
```

 

```{r}
emm.s <- emmeans(mod, specs = pairwise ~ Dx:Sex) 
emm.s
Means <- as.data.frame(emm.s$emmeans)
print(round(Means, 2))
```

 

### Scatterplot Age. vs. Predicted Age for Each Sex*Dx Group from lm(age_prediction ~ Age, data = .)
```{r,fig.width=11,height=10}
 
for (dx in c(0, 1)) {
  for (sex in c(1,2)){
    if (dx == 0) { 
      dxTitle = "HC"} else{
        dxTitle = "BD"
      }
    if (sex == 1) { 
      sexTitle = "Men"} else{
        sexTitle = "Women"
      }  
    m <- df%>%
      filter(Sex==sex, Dx==dx) %>%
      lm(age_prediction ~ Age, data = .)
    #print("BD Males")
    #print(round(summary(m)$r.squared,2))
    print(round(sqrt(summary(m)$r.squared), 2))
    
    p <- df%>%
      filter(Sex==sex, Dx==dx)%>%
      ggplot(aes(x = Age, y = age_prediction, colour = factor(Site) )) + ggtitle(paste('Age vs. Predicted Age among', dxTitle, sexTitle))+
      geom_point(size = 3, alpha = 0.75)  +
      scale_fill_manual(values = rainbow(28))+theme_classic()+labs(y = "Predicted Age") + 
      geom_smooth(fullrange = T, method = "lm", se=FALSE, color="black", formula = y~x)+ theme(legend.position = "none")+theme(axis.text=element_text(size=12),
      axis.title=element_text(size=12,face="bold"),plot.title = element_text(size = 20, face = "bold"))+ geom_abline(slope=1)+xlim(-5,90)+ylim(-5,90)                                  # Add text element to plot
      #annotate("text", x = 80, y = 40.0, label = paste0('r = ' , round(sqrt(summary(m)$r.squared), 2))) + 
     # annotate("text", x = 80, y = 45.0, label = paste(bquote('Y-axis '("a"^2))))
    print(paste0(dxTitle, sexTitle))
      jpeg(filename = paste0('figures/',dxTitle,'_',sexTitle,'BrainPAD_corr_age.jpeg'))
      print(p) 
      dev.off() 
         print(p) 
  }
} 
```


```{r,fig.width=11,height=10}
 
for (dx in c(0, 1)) {
  for (sex in c(1,2)){
    if (dx == 0) { 
      dxTitle = "HC"} else{
        dxTitle = "BD"
      }
    if (sex == 1) { 
      sexTitle = "Men"} else{
        sexTitle = "Women"
      }  
    m <- df%>%
      filter(Sex==sex, Dx==dx) %>%
      lm(age_prediction ~ Age, data = .)
    #print("BD Males")
    #print(round(summary(m)$r.squared,2))
    print(round(sqrt(summary(m)$r.squared), 2))
    
    p <- df%>%
      filter(Sex==sex, Dx==dx)%>%
      ggplot(aes(x = Age, y = age_prediction, colour = factor(Site) )) + ggtitle(paste('Age vs. Predicted Age among', dxTitle, sexTitle))+
      geom_point(size = 3, alpha = 0.75)  +
      scale_fill_manual(values = rainbow(28))+theme_classic()+labs(y = "Predicted Age") + 
      geom_smooth(fullrange = T, method = "lm", se=FALSE, color="black", formula = y~x)+ theme(legend.position = "none") +theme(axis.text=element_text(size=12),
      axis.title=element_text(size=20,face="bold"),plot.title = element_text(size = 20, face = "bold"))+ geom_abline(slope=1)+xlim(-5,90)+ylim(-5,90)+annotate("text", x = 80, y = 40.0, size = 8, label = paste0('r = ' , round(sqrt(summary(m)$r.squared), 2)))   
    print(paste0(dxTitle, sexTitle))
      jpeg(filename = paste0('figures/',dxTitle,'_',sexTitle,'BrainPAD_corr_age.jpeg'))
      print(p) 
      dev.off() 
         print(p) 
  }
} 
```

### Correlation between Age and Brain-PAD (r and scatterplot) using simple pearson  correlation
```{r}
for (dx in c(0, 1)) {
 # for (sex in c(1,2)){
    if (dx == 0) { 
      dxTitle = "HC"} else{
        dxTitle = "BD"
      }
    if (sex == 1) { 
      sexTitle = "Men"} else{
        sexTitle = "Women"
      }  
p <- df%>%
      filter(Sex==sex, Dx==dx)%>% ggplot(  aes(x = Age, y = BrainPAD, color = factor(Site))) + ggtitle(paste('Age vs. BrainPAD among', dxTitle ))+
      geom_point(size = 3, alpha = 0.75)  +
      scale_fill_manual(values = rainbow(28))+theme_classic()+labs(y = "BrainPAD") + 
      geom_smooth(fullrange = T, method = "lm", se=FALSE, color="black", formula = y~x)+ theme(legend.position = "none")+theme(axis.text=element_text(size=12),
      axis.title=element_text(size=12,face="bold"),plot.title = element_text(size = 20, face = "bold"))   +xlim(10,90)+ylim(-50,50) + geom_text(aes(x = max(Age), y = min(BrainPAD), label = paste0("r = ", round(cor(Age, BrainPAD), 2))  ), hjust = 1, vjust = 0, size = 5 , color = "black", x = 80, y = 45.0 )

print(p)
# }
}
```
### correlation between BrainPAD and Age extracted from linear model. (INCORRECT METHOD)
```{r}

# for(x in c(0,1)){
#   if(x == 0) { 
#     name = "HC"
#   } else {
#     name = "BD"
#   }
#   print(x)
#  # m <- lmer(BrainPAD ~ Age + (1 | Site), data = df, subset = Dx == x)
#  
#   mod <- lmer(BrainPAD ~ Age + Sex + (1|Site), data = df %>% filter(Dx == dx))
#   r2 <- r.squaredGLMM(mod) 
#   
#   r2c <-  r2[1, "R2c"]  # Select the marginal R-squared value
#   print(round(r2c, 2))
#   
#   r <- sqrt(r2c)
#   print(round(r, 2))
#  
#   p <- ggplot(data = df, aes(x = Age, y = BrainPAD, color = factor(Site_Locations))) +
#        geom_point(size = 2, alpha = 0.75) + 
#        scale_fill_manual(values = rainbow(28)) +
#        labs(y = "Brain-PAD") +
#        theme_classic() +
#        ggtitle(paste0('Age vs. Brain-PAD in ' , name))  +
#        geom_smooth(method = "lm", se = FALSE, color = "black", formula = y ~ x, fullrange = TRUE) +
#        theme(legend.position = "none") +
#        theme(axis.text = element_text(size = 12),
#              axis.title = element_text(size = 12, face = "bold"),
#              plot.title = element_text(size = 12, face = "bold")) +
#        xlim(0, 90) +
#        ylim(-40, 40) +
#        annotate("text", x = 80, y = 40.0, label = paste0("r = -", round(r, 3) ))
# 
#   #jpeg(filename = paste0('figures/',name,'BrainPAD_corr_age.jpeg'))
#   print(p) 
#   #dev.off()
# }

```


### BrainPAD by Age Group
```{r}

for( x in c(0,1)){
  if (x == 1) { 
  name = "BD"} else{
    name = "HC"
  }
p<- df %>%
  mutate(AgeGroup = case_when(
    (Age >= 18 & Age <= 29)~"18-29",
   # (Age >= 20 & Age <= 29)~ "25-29",
    (Age >= 30 & Age <= 39)~ "30-39",
    (Age >= 40 & Age <= 49)~ "40-49",
    (Age >= 50 & Age <= 59)~ "50-59",
    (Age >= 60 & Age <= 75)~"60-75" 
  ))   %>% filter(Dx==x, is.na(AgeGroup) == FALSE ) %>%
  ggplot(mapping = aes(x=AgeGroup, y = BrainPAD, fill=factor(AgeGroup))) + geom_violin(alpha=0.25)+
  scale_fill_manual(values = rainbow(5)) + theme_classic() + guides(fill=guide_legend(title="Age Group")) + ggtitle(paste("BrainPAD across Age Groups in",name)) + theme(
  axis.title.x = element_text(size = 20),
 plot.title = element_text(size = 20, face = "bold"),
 axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18,  hjust = 1, vjust = 0.5),
 axis.text.y=element_text(size=18))  
print(paste0('figures/',name,'BrainPAD_across_age.jpeg'))
jpeg(filename = paste0('figures/',name,'BrainPAD_across_age.jpeg'))
print(p) 
dev.off()
  
}
 
```
 