---
title: "photon_regression_models"
author: "Hui Xin Ng"
date: "2022-04-15"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
library("dplyr")
library("tidyverse")
library("DescTools")
library("ggplot2")
library("readxl")
library("lme4")
library("ModelMetrics")
library("merTools")
library("lmtest")
library("emmeans")
library("haven")
library("lmerTest")
library("metaforest")
library("rstatix")
library("ech")
library("labelled")
 
libs <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest", "metaforest", "rstatix",  'ech', 'labelled')
#require(libs)
lapply(libs, require, character.only = TRUE)
```

```{r}
replaceSpace <- function(dataframe) {
   gsub(x = names(dataframe), pattern = " ", replacement = "_") 
} 

## Helper functions
getPartialEtaSq<-function(m_fixed) {
  petasq <- round(as.data.frame(EtaSq(m_fixed))[2],3)
  return(petasq) }

getClinVarCoeff <- function(m) { # get coefficients
  as.data.frame(summary(m)$coefficients)[6, ]
} 
path = file.path( "Master_BD_080720_withPredictions_withNewHajek_fixed.sav")
df = read_sav(path)

# Helper function to rename df columns
replaceSpace <- function(dataframe) {
  gsub(x = names(dataframe), pattern = " ", replacement = "_") 
}

df <- df %>% 
  mutate_if(is.character, list(~na_if(.,""))) %>%
  filter(!SubjID %in%  c('con_11',  'pat_13')) %>% # drop 2 subjects from Haarman
  arrange(Site) # arrange by Site number

siteLoc <- read_excel("DescriptiveData_080720_2.xlsx", sheet=3)

names(siteLoc) <- replaceSpace(siteLoc)
names(df) <- replaceSpace(df) 

df <- merge(df, siteLoc, by=c("PIName","Site" )) # merge 
df <- df %>% 
filter(Age <= 75  & Age  >=18)  
BD_only <- df %>% filter(Dx==1)
BD_only
```

```{r}
mod <- lmer(BrainPAD ~ Dx + Sex + Age + (1|Site) , data = df)
 
# Calculate the correlation between the residuals of the model and x and y
# no correlation between residuals of the model including Age and Age itself
resid <- residuals(mod)
cor.test(resid, df$Age) 
```
 
```{r}
 
# Create a scatterplot of Brain-PAD vs. Age 
plot(df$Age, df$BrainPAD,  xlab = "Age", ylab = "Brain-PAD", )
 

# Create a scatterplot of residuals vs. Age
plot(df$Age, residuals(mod), xlab = "Age", ylab = "Brain-PAD Corrected for Age")
abline(h = 0, lty = 2)

``` 

```{r} 
 
modAge2 <- lmer(BrainPAD ~ Dx + Sex +   Age + Age2 + (1 | Site) , data = df) 
summary(modAge2)

# Calculate the correlation between the residuals of the model and x and y
# no correlation between residuals 
resid <- residuals(modAge2)
cor.test(resid, df$Age)


plot(df$Age, residuals(modAge2), xlab = "Age",   ylab = bquote("Brain-PAD corrected for Age & "~Age^2))
# Add a horizontal regression line to the scatterplot
abline(h = 0, lty = 2)
```
### Determine if we need to control for any non linear age effects

```{r}
# Supplementary table 1 
modAge2Age3 <- lmer(BrainPAD ~ Dx + Sex +   Age + Age2 + Age3 + (1 | Site) , data = df) 
summary(modAge2Age3)

# Calculate the correlation between the residuals of the model and x and y
resid <- residuals(modAge2Age3)
cor.test(resid, df$Age)
```
 
```{r}
anova(modAge2, modAge2Age3, test= "chisq")
```

```{r}
anova(modAge2, mod, test= "chisq")
```

### EMMs of  Dx extracted from first basic model
```{r}
# estimated marginal means for Dx group
emm.s <- emmeans(mod, specs = pairwise ~ Dx) 
Means = round(as.data.frame(emm.s$emmeans), 2)
Means
# The EMM of BrainPADs is -0.95 (SE 0.68, CI: -2.34 - 0.44), for HC and
# +1.42  years (SE 0.66,  CI: 0.05 - 2.78) for BD individuals. 

jpeg(filename = paste0('figures/', 'compare_emmeans_HC_BD.jpeg'), width=600, height=600)
p <- plot(emm.s, comparisons = TRUE) + theme_classic() + scale_y_discrete("Groups", labels = c("HC",'BD')) + scale_x_continuous("Brain-PAD") +coord_flip()   + theme(
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
 axis.text.x=element_text(size=18),
 axis.text.y=element_text(size=18)) #+ggtitle("Mean estimate across groups and 95% CI of BrainPAD")  

# Get the counts for each medication group
counts <- df %>% 
  filter(!is.na(Dx)) %>%
  group_by(Dx) %>% 
  summarize(count = n())
 
dev.off()
```



 
### Model accounting for interaction between Dx and Age, Dx and Sex, and Age and Sex, and three way interaction between Dx Age and Sex
```{r}

modDxAgeSexFull <- lmer(BrainPAD ~ Dx + Sex +  Age  + (Dx*Age) + (Dx*Sex) + (Age*Sex) + (Dx*Age*Sex)+  (1 | Site) , data = df) 
summary(modDxAgeSexFull)

m1_fixed <- lm(BrainPAD ~ Dx + Sex +  Age  + (Dx*Age) + (Dx*Sex) + (Age*Sex) + (Dx*Age*Sex)+  + Site, data =df)
summary(m1_fixed)
print(round(summary(m1_fixed)$r.squared,2))
# Use anova to get f values 
getPartialEtaSq(m1_fixed)

```

```{r}
modDxAgeSexNoThreeWay <- lmer(BrainPAD ~ Dx + Sex +  Age  + (Dx*Age) + (Dx*Sex) + (Age*Sex) +    (1 | Site) , data = df) 
summary(modDxAgeSexNoThreeWay)

m1_fixed <- lm(BrainPAD ~ Dx + Sex +  Age  + (Dx*Age) + (Dx*Sex) + (Age*Sex) +    Site, data =df)
summary(m1_fixed)
print(round(summary(m1_fixed)$r.squared,2))
# Use anova to get f values 
getPartialEtaSq(m1_fixed)  
```

 
# get correlation between brain age and age for BD and HC FROM LMM MODEL
R-squared is a measure of how much of the variability in the response variable is explained by the predictor variables in the model.   
Sqrt of R-squared is the correlation between the observed and predicted values of the response variable.
```{r}

# # Correlations between age and brain-PAD for HC and BD are 0.64  and 0.53respectively
# for (sex in c(1, 2)) {
# for (dx in c(0, 1)) {
#   print(dx)
#   mod <- lmer(BrainPAD ~ Age + Sex + (1|Site), data = df %>% filter(Dx == dx))
#     r2 <- r.squaredGLMM(mod) 
#     r2c <-  r2[1, "R2c"]  # Select the marginal R-squared value 
#     r <- sqrt(r2c)
#     print(round(r, 2))
#     if (dx == 0) {
#       dxTitle = "HC"} else{
#         dxTitle = "BD"
#       }
#     
#     p <- df%>% filter(Sex==sex, Dx==dx)%>%
#       ggplot(aes(x = Age, y = BrainPAD, colour = factor(Site) )) + ggtitle(paste('Age vs. Brain-PAD among', dxTitle ))+
#       geom_point(size = 3, alpha = 0.75)  +
#       scale_fill_manual(values = rainbow(28))+theme_classic()+labs(y = "Brain-PAD") + geom_smooth(fullrange = T, method = "lm", se=FALSE, color="black", formula = y~x)+
#       theme(legend.position = "none")+theme(axis.text=element_text(size=12), axis.title=element_text(size=12,face="bold"),plot.title = element_text(size = 12, face = "bold")) + annotate("text", x = 80, y = 40.0, label = paste0('r = ' , round(r, 2)))
# 
# jpeg(filename = paste0('figures/',dxTitle,'_' ,'BrainPAD_corr_age.jpeg', width = 1200, height = 1200,  res = 200 ))
# print(p)
# dev.off()
#   }  
# }
```
 
# Stepwise removal of interaction terms from three way interaction model
```{r}
modNoDxSex <- lmer(BrainPAD ~ Dx + Sex + Age + (Dx*Age) + (Age*Sex) +(1 | Site) , data = df)  
summary(modNoDxSex)

# modNoAgeSex <- lmer(BrainPAD ~ Dx + Sex + Age + (Dx*Age) + (Dx*Sex) + (1 | Site) , data = df)  
# modNoDxAge <- lmer(BrainPAD ~ Dx + Sex + Age + (Age*Sex) + (Dx*Sex) + (1 | Site) , data = df)  

fixed_modNoDxSex <- lm(BrainPAD ~ Dx + Sex + Age + (Dx*Age) + (Age*Sex) + Site, data = df)
summary(fixed_modNoDxSex)
getPartialEtaSq(fixed_modNoDxSex)

```


### Model account for interaction between  Dx and Sex
```{r}
modDxSex <- lmer(BrainPAD ~ Dx + Age + Sex  +  Dx*Sex   + (1 | Site), data = df) 
summary(modDxSex)

#linear model with only fixed effects for Dx*Sex
m1_fixed <- lm(BrainPAD ~    Dx+ Age + Sex  + Dx*Sex + Site, data =df)
summary(m1_fixed)
print(round(summary(m1_fixed)$r.squared,2))

# Use anova to get f values 
getPartialEtaSq(m1_fixed) 
```

### Model account for interaction between  Sex and Age
Which sex has higher/lower brain age - EMM 
```{r}
modSexAge <- lmer(BrainPAD ~ Dx + Age + Sex  +  Sex*Age   + (1 | Site), data = df) 
summary(modSexAge) 

#linear model with only fixed effects for Dx*Sex
modSexAge_fixed <- lm(BrainPAD ~    Dx+ Age + Sex  +  Sex*Age + Site , data =df)
summary(modSexAge_fixed)
print(round(summary(modSexAge_fixed)$r.squared,2))
# Use anova to get f values 
getPartialEtaSq(modSexAge_fixed)

```


### Model account for interaction between  Dx and Age
```{r}
modDxAge <- lmer(BrainPAD ~ Dx + Age + Sex   +  Dx*Age   + (1 | Site), data = df) 
summary(modDxAge)
 
#linear model with only fixed effects 
modDxAge_fixed <- lm(BrainPAD ~    Dx+ Age + Sex  +  Dx*Age + Site , data =df)
summary(modDxAge_fixed)
print(round(summary(modDxAge_fixed)$r.squared,2))
# Use anova to get f values 
getPartialEtaSq(modDxAge_fixed)

```

```{r} 
modDxAge <- lmer(BrainPAD ~ Dx + Sex +  Age  + (Dx*Age) + (1 | Site) , data = df) # remvove Sex*Age and Dx*Sex
anova(modDxAgeSexNoThreeWay, modDxAgeSexFull, test= "chisq")    
anova(modNoDxSex, modDxAgeSexNoThreeWay, test= "chisq") # remove Dx*Sex 
```

### EMMs of  Dx extracted from the final model after stepwise selection
```{r}
# estimated marginal means for Dx group
emm.s <- emmeans(modNoDxSex, specs = pairwise ~ Dx) 
Means = round(as.data.frame(emm.s$emmeans), 2)
Means
# The EMM of BrainPADs is -0.95 (SE 0.68, CI: -2.34 - 0.44), for HC and
# +1.42  years (SE 0.66,  CI: 0.05 - 2.78) for BD individuals.  
# Create a data frame for plotting
plot_data <- as.data.frame(emm.s$emmeans)

# Define custom order for x-axis
custom_order <- c('HC', 'BD')

jpeg(filename = paste0('figures/', 'compare_emmeans_HC_BD.jpeg'), width=600, height=600)

# Create the ggplot with a similar visual style
 ggplot(plot_data, aes(x = emmean, y = Dx, xmin = lower.CL, xmax = upper.CL)) +
  geom_vline(xintercept = 0, color = "gray70", linetype = "dashed", size = 0.5) +
  geom_point(size = 4, color = "red") +
  geom_errorbarh(height = 0.15) +
  labs(x = "Brain-PAD (years)", y = "") +
  scale_x_continuous(limits = c(-3, 3), breaks = seq(-3, 3, by = 1), expand = c(0.01, 0.01)) +
  scale_y_discrete(limits = custom_order, labels = c('HC\nN=1342', 'BD\nN=1577')) +
  theme_minimal() 
dev.off() 
```
 
 

## BD ONLY ANALYSIS
### Full Model with AntiEpileptic, Li, BDI and Gen2AntiPsychotics
```{r}
m1 <- lmer(BrainPAD ~    Age + Sex  + AntiEpileptic + Li+BDI+ Gen2AntiPsych + (1 | Site), data =BD_only) 
# Table 1 in paper: Model with all significant clinical variables within BD only
summary(m1)  

BD_only
```

```{r}
#linear model with only fixed effects
m1_fixed <- lm(BrainPAD ~    Age + Sex  + AntiEpileptic + Li+BDI+ Gen2AntiPsych + Site, data =BD_only)
summary(m1_fixed)
print(round(summary(m1_fixed)$r.squared,2))
# Use anova to get f values 
getPartialEtaSq(m1_fixed)
``` 

### Fit each clinical variable respectively with the baseline model 
```{r}
# Bipolar I
mod <- lmer(BrainPAD ~   Age + Sex  + BDI + (1 | Site)  , data = BD_only)   
summary(mod)  
# 11.74/(11.74 + 68.72)
emm.s <- emmeans(mod, specs = pairwise ~ BDI)
round(as.data.frame(emm.s$emmeans), 2) 
```

```{r}
mod <- lmer(BrainPAD ~   Age + Sex  + BDI + (1 | Site)
           , data = BD_only)
summary(mod) 
emm.s <- emmeans(mod, specs = pairwise ~ BDI)
print(round(as.data.frame(emm.s$emmeans), 2))
pvalBDI<- getClinVarCoeff(mod)

mod_fixed <- lm(BrainPAD ~   Sex + Age   + BDI + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```

```{r}
mod <- lmer(BrainPAD ~   Age + Sex  + Li + (1 | Site)
           , data = BD_only)
summary(mod)
emm.s <- emmeans(mod, specs = pairwise ~ Li)
print(round(as.data.frame(emm.s$emmeans), 2))
pvalLi<- getClinVarCoeff(mod)

```

```{r }
mod_fixed <- lm(BrainPAD ~    Sex + Age  + Li + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed)  
mod_fixed

``` 

```{r}
mod <- lmer(BrainPAD ~    Age + Sex   + AntiEpileptic + (1 | Site)
           , data =BD_only) 
summary(mod) # p-values and beta coeff 
```

```{r}
AntiE <- getClinVarCoeff(mod)

emm.s <- emmeans(mod, specs = pairwise ~ AntiEpileptic)
round(as.data.frame(emm.s$emmeans), 2) # EMMs
  
pvalAntiEpileptic <- getClinVarCoeff(mod) 
```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age   + AntiEpileptic + Site
           , data = BD_only) 
getPartialEtaSq(mod_fixed) 

```
 
```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Gen2AntiPsych + (1 | Site)
           , data = BD_only)  
summary(mod) 
emm.s <- emmeans(mod, specs = pairwise ~ Gen2AntiPsych)
round(as.data.frame(emm.s$emmeans), 2)
pvalGen2AntiPsych <- getClinVarCoeff(mod)

```

```{r}
mod_fixed <- lm(BrainPAD ~    Sex + Age   + Gen2AntiPsych + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~   Age + Sex  + Gen1AntiPsych + (1 | Site) , data = BD_only)  
summary(mod)
emm.s <- emmeans(mod, specs = pairwise ~ Gen1AntiPsych)
round(as.data.frame(emm.s$emmeans), 2)
  
pvalGen1AntiPsych <- getClinVarCoeff(mod)
```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age  + Gen1AntiPsych + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~   Age + Sex  + AntiDep + (1 | Site), data =  BD_only) 
summary(mod)  
emm.s <- emmeans(mod, specs = pairwise ~ AntiDep)
round(as.data.frame(emm.s$emmeans), 2)
 
pvalAntiDep <- getClinVarCoeff(mod)
 
```

```{r}
mod_fixed <- lm(BrainPAD ~    Sex + Age   + AntiDep + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
mod_fixed
```


```{r}
# (b = -0.09, p = 0.90, ηp² = 0.007)
mod <- lmer(BrainPAD ~  Age + Sex + Euthymic + (1 | Site)
           , data= BD_only) 
summary(mod) 

emm.s <- emmeans(mod, specs = pairwise ~ Euthymic)
round(as.data.frame(emm.s$emmeans), 2) 

 


pvalEuthymic <- getClinVarCoeff(mod)

```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age  + Euthymic + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~   Age + Sex  + HistoryPsychosis + (1 | Site)
           , data =BD_only)  
emm.s <- emmeans(mod, specs = pairwise ~ HistoryPsychosis)
round(as.data.frame(emm.s$emmeans), 2)
 

pvalHistoryPsychosis <- getClinVarCoeff(mod)
pvalHistoryPsychosis
```
 
```{r} 
mod_fixed <- lm(BrainPAD ~   Sex +Age + HistoryPsychosis + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```
 

```{r}
mod <- lmer(BrainPAD ~   Age + Sex   + AgeofOnset + (1 | Site)
           , data =  BD_only) 
pvalAgeofOnset <- getClinVarCoeff(mod)
pvalAgeofOnset
summary(mod)
```

```{r}
mod_fixed <- lm(BrainPAD ~   Age + Sex  + AgeofOnset + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed)
```
### BH corrected p-values to adjust for false discovery rates
```{r}
pvalues_df <- rbind(pvalBDI, pvalLi, pvalAntiEpileptic,  pvalGen2AntiPsych, pvalGen1AntiPsych, pvalEuthymic, pvalAntiDep ,  pvalHistoryPsychosis,  pvalAgeofOnset	)
pvalues <- pvalues_df$`Pr(>|t|)` 
print(pvalues)
round(pvalues, 2)
print(p.adjust(pvalues,method="BH"))  
round(p.adjust(pvalues,method="BH"), 4)

pvalues
# pvalBDI, pvalLi, pvalAntiEpileptic,  pvalGen2AntiPsych,
```

### Model with Mood State collapsed into three levels 
```{r}
BD_only <- df %>% filter(Dx==1) 
BD_only <-unlabelled(BD_only) %>% group_by(MoodState) %>% dplyr::select(BrainPAD, MoodState, Age, Age2, Age3, Sex, Site) %>% na.omit()  #%>% tally() 

BD_only <-  (BD_only) %>% mutate(EuthDepOther =
                  case_when(
                    MoodState == "Euthymic"  ~ 0,
                    MoodState == "Depressed"  ~ 1,
                    MoodState == "Manic"    ~ 2,
                    MoodState == "Mixed" ~ 2,
                    MoodState == "Hypomanic" ~2 ))  %>% mutate(EuthDepOther = as.factor(EuthDepOther)) 

BD_only %>% group_by(EuthDepOther)%>% tally() 


```


### Violin plots Mood State collapsed into three levels 
```{r}

brainPADs <- BD_only%>% 
  filter( !is.na(EuthDepOther)) %>%
  group_by(EuthDepOther) %>% 
  dplyr::summarize( brainPAD_mean = round(mean(BrainPAD), 2),
                    , count = sum(EuthDepOther))  

jpeg(filename = paste0('figures/', 'MoodStateCollapsed.jpeg'), width=600, height=600)
BD_only %>% 
  dplyr::select(BrainPAD, EuthDepOther) %>% 
   reshape2::melt(id.vars = c("EuthDepOther", "BrainPAD")) %>% 
  na.omit() %>% 
  ggplot(mapping = aes(x = factor(EuthDepOther), y = BrainPAD, fill = factor(EuthDepOther))) +
  ylim(-45, 45) +
  geom_violin(alpha = 0.3) +
  theme_classic()  +stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "red", size = 0.25)  + theme(
  axis.title.x = element_text(size = 20),
  plot.title = element_text(size = 20, face = "bold"),
  axis.title.y = element_text(size = 20),
  axis.text.x=element_text(size=18),
  axis.text.y=element_text(size=18))   + theme(legend.position = "none") +
  scale_x_discrete(labels = c("Euthymic", "Depressed", "Any Mania"))  + xlab("Mood State") + 
  geom_label(data = brainPADs, aes(x = factor(EuthDepOther), y = brainPAD_mean, label = paste(brainPAD_mean, "\nN=", count)), alpha= 0.2, nudge_x = 0, nudge_y = 40, size = 6) 
dev.off()
```

```{r} 
mod_EuthDepOther <- lmer(BrainPAD ~   Age + Sex  +  EuthDepOther  + (1 | Site) , data = BD_only )
#summary(mod_EuthDepOther)
# Calculate emmeans
emmeans_obj <- emmeans(mod_EuthDepOther, specs = pairwise ~"EuthDepOther")

(emmeans_obj)
 


```
```{r}
mod_fixed <- lm(BrainPAD ~   Age + Sex  + EuthDepOther + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed)   
```
 
# PredictInterval is not relevant for us
```{r}
# Prediction of response values using new data, based on distributions of params of the model
# uses arm::sim function
#predictInterval(mod)   # for various model predictions, possibly with new data
```
```{r}
# NOT RELEVANT SINCE DX*Sex is not sig
# ### EMMs of Sex by Dx extracted from first basic model
# perform pairwise comparisons between the 4 different combinations of sex and diagnosis using the EMMs
# See table X for the EMMs for each group
# emm.s <- emmeans(mod, specs = pairwise ~ Dx:Sex) 
# Means = round(as.data.frame(emm.s$emmeans), 2)
# 
# # Men (+2.37 years; p <0.0001) and women (+2.37 years; p <0.0001) with BD had significantly higher EMM of brainPAD compared to their healthy counterparts

# effects associated with Site are uncorrelated with the means of the fixed effects (Dx, Age)
# residual plot
# plot(mod, resid(., type = "pearson") ~ fitted(.),
#      abline = 0, pch = 20, cex = .8, id = 0.05,
#      xlab = "Ftted values", ylab = "Pearson Residuals")
 
``` 
 

```{r} 
df$DurationIll = df$Age-df$AgeofOnset
ggplot(df, aes(x = AgeofOnset)) + 
  geom_histogram(aes(y = ..density..), binwidth = 2, color = "black", fill = "white") + 
  geom_density(alpha = .2, fill = "#FF6666") + 
  ggtitle("Distribution of X") + 
  xlab("X") + theme_classic()
  ylab("Density")   
  
df %>% filter(!is.na(AgeofOnset)) %>%  filter( 0< (AgeofOnset))  %>% 
  summarise(mean = mean(AgeofOnset), sd = sd(AgeofOnset), min = min(AgeofOnset))
 
df %>% filter(!is.na(DurationIll)) %>%# 22.27, 9
  summarise(mean = mean(DurationIll), sd = sd(DurationIll)) 
 

ggplot(df, aes(x = DurationIll)) + 
  geom_histogram(aes(y = ..density..), binwidth = 2, color = "black", fill = "white") + 
  geom_density(alpha = .2, fill = "#FF6666") + 
  ggtitle("Distribution of X") + 
  xlab("X") + theme_classic()
  ylab("Density") 
```