---
title: "photon_regression_models"
author: "Hui Xin Ng"
date: '2022-04-15'
output: html_document
---
```{r setup, include=FALSE}
library("dplyr")
library("tidyverse")
library("DescTools")
library("ggplot2")
library("readxl")
library("lme4")
library("ModelMetrics")
library("merTools")
library("lmtest")
library("emmeans")
library("haven")
library("lmerTest")
library("metaforest")
library("rstatix")
library("ech")
library("labelled")
 
libs <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest", "metaforest", "rstatix",  'ech', 'labelled')
#require(libs)
lapply(libs, require, character.only = TRUE)
replaceSpace <- function(dataframe) {
   gsub(x = names(dataframe), pattern = " ", replacement = "_") 
} 

## Helper functions
getPartialEtaSq<-function(m_fixed) {
  petasq <- round(as.data.frame(EtaSq(m_fixed))[2],3)
  return(petasq)
}

getClinVarCoeff <- function(m) {
  as.data.frame(summary(m)$coefficients)[6, ]
}

plotRandEff <- function(m) {
  REsim_df <- REsim(m)
names(siteLoc)[3] <- "groupID"
REsim_sitename <- merge(REsim_df, siteLoc, by=c( "groupID"))

REsim_sitename$groupFctr <- "" 
REsim_sitename$groupID <- REsim_sitename$Site_Locations
plotREsim(REsim_sitename, labs = TRUE) + xlab('') +coord_flip() + 
  ggtitle("Simulation of Random Effects Confidence Intervals") 
}

path = file.path("Master_BD_080720_withPredictions_withNewHajek_fixed.sav")
df = read_sav(path)
siteLoc <- read_excel("DescriptiveData_080720_2.xlsx", sheet=3)

df <- df %>% # create AgeGroup var
  mutate(AgeGroup = case_when(
    (Age >= 18 & Age <= 29)~"18-29",
   # (Age >= 20 & Age <= 29)~ "25-29",
    (Age >= 30 & Age <= 39)~ "30-39",
    (Age >= 40 & Age <= 49)~ "40-49",
    (Age >= 50 & Age <= 59)~ "50-59",
    (Age >= 60 & Age <= 75)~"60-75" 
  ))  

df <- df %>% # remove two subjects from Hajek
  mutate_if(is.character, list(~na_if(.,""))) %>%
  filter(!SubjID %in%  c('con_11',  'pat_13')) %>%
    arrange(Site) 

names(siteLoc) <- replaceSpace(siteLoc)
names(df) <- replaceSpace(df) 

df <- merge(df, siteLoc, by=c("PIName","Site"  )) # add these columns to the df
BD_only <- df %>% filter(Dx==1) # create BD only df


```
 

 
### First model with linear and non linear ages, with Sex and Dx  
This model is used to extract est. marginal means for both Sex*Dx and Dx Groups
```{r} 
mod <- lmer(BrainPAD ~ Dx + Sex +   Age + Age2 + Age3 + (1 | Site) , data = df) 
summary(mod)
```

# estimated random effect for each site and their interval estimate
```{r}
# get some sense of uncertainty of the random effect of each site

plotRandEff(mod) 

impSim <- REimpact(mod, df[7, ], groupFctr = "Site", breaks = 5,
                   n.sims = 300, level = 0.9)
ggplot(impSim, aes(x = factor(bin), y = AvgFit, ymin = AvgFit - 1.96*AvgFitSE,
                   ymax = AvgFit + 1.96*AvgFitSE)) +
  geom_pointrange() + theme_bw() + labs(x = "Bin of `d` term", y = "Predicted Fit")
```
### EMMs of  Dx extracted from first basic model
```{r}
# estimated marginal means for Dx group
emm.s <- emmeans(mod, specs = pairwise ~ Dx) 
Means = round(as.data.frame(emm.s$emmeans), 2)
Means
# The EMM of BrainPADs is -0.95 (SE 0.68, CI: -2.34 - 0.44), for HC and
# +1.42  years (SE 0.66,  CI: 0.05 - 2.78) for BD individuals. 

plot(emm.s, comparisons = TRUE) + theme_classic() + scale_y_discrete("Groups", labels = c("HC",'BD')) + scale_x_continuous("Est. Marginal Means") +coord_flip()
 
```

### Basic Model with site as fixed effect to extract eta-squared
```{r}
mod_fixed <- lm(BrainPAD ~ Dx + Sex +  Age + Age2 + Age3 + Site, data = df)   
 
#round(as.data.frame(anova(mod_fixed)),3) # Use anova to get f values

summary(mod_fixed)
# The BD group showed +2.37 (SE 0.31) years higher brain-PAD than controls (F(2, 2193) =  127.3, p<0.0001) when adjusted for age, age2, age3, sex, and scanning site 

EtaSq(mod_fixed)  
# Partial eta-squared with p-values was calculated for each of the fixed effects based on the F-statistic 
```
 
### Model accounting for interaction between Dx and Age, Dx and Sex, and Age and Sex, and three way interaction between Dx Age and Sex
```{r}
mod <- lmer(BrainPAD ~ Dx + Sex +  Age + Age2 + Age3 + (Dx*Age) + (Dx*Sex) + (Age*Sex) + (Dx*Age*Sex)+  (1 | Site) , data = df) 
summary(mod)
# ICC - how much of the variation in the response variable, which is not attributed to fixed effects, is accounted for by a random effect.
# https://www.ssc.wisc.edu/sscc/pubs/MM/MM_DiagInfer.html
# 11.24/(11.24 + 61.98)
```

 
# get correlation between brain age and age for BD and HC 
```{r}
# Correlations between age and brain-PAD for HC and BD are -0.49 and -0.41 respectively 
for (dx in c(0,1)){
  mod <- df %>% filter(Dx == dx) %>% 
  lm(data = ., Age ~ BrainPAD) %>% summary()  
  r <- round(sqrt(mod$r.squared),2)
  print(r)
  sqrtOfr2 <- df %>% filter(Dx == dx)  %>% 
  summarise(out = cor(Age, BrainPAD, use = "complete.obs")) %>% round(2)
  print(sqrtOfr2)
  
}

```




### Model account for interaction between  Dx and Sex
```{r}
mod <- lmer(BrainPAD ~ Dx + Age + Sex + Age2 + Age3 +  Dx*Sex   + (1 | Site), data = df) 
summary(mod)
```

### Model account for interaction between  Dx and Age
```{r}
mod <- lmer(BrainPAD ~ Dx + Age + Sex + Age2 + Age3 +  Dx*Age   + (1 | Site), data = df) 
summary(mod)
```

### Model account for interaction between  Sex and Age
```{r}
mod <- lmer(BrainPAD ~ Dx + Age + Sex + Age2 + Age3 +  Sex*Age   + (1 | Site), data = df) 
summary(mod)
```

## BD ONLY ANALYSIS
### Full Model with AntiEpileptic, Li, BDI and Gen2AntiPsychotics
```{r}
m1 <- lmer(BrainPAD ~    Age + Sex + Age2 + Age3 + AntiEpileptic + Li+BDI+ Gen2AntiPsych + (1 | Site), data =BD_only) 

summary(m1) 

plotRandEff(m1)
```

```{r}
#linear model with only fixed effects
m1_fixed <- lm(BrainPAD ~    Age + Sex + Age2 + Age3 + AntiEpileptic + Li+BDI+ Gen2AntiPsych + Site
           , data =BD_only)
summary(m1_fixed)
# Use anova to get f values 
getPartialEtaSq(m1_fixed)
``` 

### Fit each clinical variable respectively with the baseline model 
```{r}
# Bipolar I
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + BDI + (1 | Site)
           , data = BD_only)   
summary(mod)  
# 11.74/(11.74 + 68.72)
emm.s <- emmeans(mod, specs = pairwise ~ BDI)
round(as.data.frame(emm.s$emmeans), 2)

plotRandEff(mod)
pvalBDI <- getClinVarCoeff(mod)

```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age + Age2 + Age3 + BDI + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 

```

```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + Li + (1 | Site)
           , data = BD_only)    
emm.s <- emmeans(mod, specs = pairwise ~ Li)
print(round(as.data.frame(emm.s$emmeans), 2))

plotRandEff(mod)

```

```{r }
mod_fixed <- lm(BrainPAD ~    Sex + Age + Age2 + Age3 + Li + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed)  
``` 

```{r}
mod <- lmer(BrainPAD ~    Age + Sex + Age2 + Age3 + AntiEpileptic + (1 | Site)
           , data =BD_only) 
summary(mod) # p-values and beta coeff 

AntiE <- getClinVarCoeff(mod)

emm.s <- emmeans(mod, specs = pairwise ~ AntiEpileptic)
round(as.data.frame(emm.s$emmeans), 2) # EMMs

plotRandEff(mod)  
pvalAntiEpileptic <- getClinVarCoeff(mod) 
```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age + Age2 + Age3 + AntiEpileptic + Site
           , data = BD_only) 
getPartialEtaSq(mod_fixed) 

```
 
```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + Gen2AntiPsych + (1 | Site)
           , data = BD_only)  
summary(mod) 
emm.s <- emmeans(mod, specs = pairwise ~ Gen2AntiPsych)
round(as.data.frame(emm.s$emmeans), 2)
pvalGen2AntiPsych <- getClinVarCoeff(mod)

```

```{r}
mod_fixed <- lm(BrainPAD ~    Sex + Age +Age2 + Age3 + Gen2AntiPsych + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + Gen1AntiPsych + (1 | Site) , data = BD_only)  
summary(mod)
emm.s <- emmeans(mod, specs = pairwise ~ Gen1AntiPsych)
round(as.data.frame(emm.s$emmeans), 2)

plotRandEff(mod) 
pvalGen1AntiPsych <- getClinVarCoeff(mod)
```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age + Age2 + Age3 + Gen1AntiPsych + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + AntiDep + (1 | Site), data =  BD_only) 
summary(mod)
emm.s <- emmeans(mod, specs = pairwise ~ AntiDep)
round(as.data.frame(emm.s$emmeans), 2)

plotRandEff(mod) 
pvalAntiDep <- getClinVarCoeff(mod)
```

```{r}
mod_fixed <- lm(BrainPAD ~    Sex + Age +Age2 + Age3 + AntiDep + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~  Age + Sex + Age2 + Age3 + Euthymic + (1 | Site)
           , data= BD_only) 
summary(mod) 

emm.s <- emmeans(mod, specs = pairwise ~ Euthymic)
round(as.data.frame(emm.s$emmeans), 2) 

plotRandEff(mod)


pvalEuthymic <- getClinVarCoeff(mod)

```

```{r}
mod_fixed <- lm(BrainPAD ~   Sex + Age + Age2 + Age3 + Euthymic + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```


```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + HistoryPsychosis + (1 | Site)
           , data =BD_only)  
emm.s <- emmeans(mod, specs = pairwise ~ HistoryPsychosis)
round(as.data.frame(emm.s$emmeans), 2)

plotRandEff(mod)

pvalHistoryPsychosis <- getClinVarCoeff(mod)

```
 
```{r}
  
mod_fixed <- lm(BrainPAD ~   Sex +Age +  Age2 + Age3 + HistoryPsychosis + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed) 
```
 

```{r}
mod <- lmer(BrainPAD ~   Age + Sex + Age2 + Age3 + AgeofOnset + (1 | Site)
           , data =  BD_only) 
plotRandEff(mod)
pvalAgeofOnset <- getClinVarCoeff(mod)

```

```{r}
mod_fixed <- lm(BrainPAD ~   Age + Sex + Age2 + Age3 + AgeofOnset + Site
           , data = BD_only)
getPartialEtaSq(mod_fixed)
```
### BH corrected p-values to adjust for false discovery rates
```{r}
pvalues_df <- rbind(pvalBDI, pvalLi, pvalAntiEpileptic,  pvalGen2AntiPsych, pvalGen1AntiPsych, pvalEuthymic, pvalAntiDep ,  pvalHistoryPsychosis,  pvalAgeofOnset	)
pvalues <- pvalues_df$`Pr(>|t|)` 
print(pvalues)
round(pvalues, 2)
print(p.adjust(pvalues,method="BH"))  
round(p.adjust(pvalues,method="BH"), 2)
```

# ARCHIVED ANALYSES
```{r}
# emm.s <-  emmeans(mod, specs = ~pairwise ~ Dx*AgeGroup )
# plot(emm.s, comparisons = TRUE) + theme_classic() + scale_y_discrete("Groups", labels = c("HC",'BD')) + scale_x_continuous("Est. Marginal Means") +coord_flip()
```
# PredictInterval is not relevant for us
```{r}
# Prediction of response values using new data, based on distributions of params of the model
# uses arm::sim function
#predictInterval(mod)   # for various model predictions, possibly with new data
```
```{r}
# NOT RELEVANT SINCE DX*Sex is not sig
# ### EMMs of Sex by Dx extracted from first basic model
# perform pairwise comparisons between the 4 different combinations of sex and diagnosis using the EMMs
# See table X for the EMMs for each group
# emm.s <- emmeans(mod, specs = pairwise ~ Dx:Sex) 
# Means = round(as.data.frame(emm.s$emmeans), 2)
# 
# # Men (+2.37 years; p <0.0001) and women (+2.37 years; p <0.0001) with BD had significantly higher EMM of brainPAD compared to their healthy counterparts

# effects associated with Site are uncorrelated with the means of the fixed effects (Dx, Age)
# residual plot
# plot(mod, resid(., type = "pearson") ~ fitted(.),
#      abline = 0, pch = 20, cex = .8, id = 0.05,
#      xlab = "Ftted values", ylab = "Pearson Residuals")
 
``` 
### Model accounting for interaction between Dx and Age
```{r}
# mod <- lmer(BrainPAD ~ Dx + Age + Sex + Age2 + Age3 + Dx*Age  + (1 | Site)
#            , data = df)  
# summary(mod)
```